Curve Average \\ {\large Project 3 for class CS6491 Computer Graphics}

### 2015-11-01 paper/paper.tex
proofread introduction chapter and visualization chapter

\section{Objective}
Formally, for two curves $A = \{a_0 ... a_{n-1}\}$ and $B = \{b_0 ... b_{n-1}\}$, each with $n$ points and $\forall x \in (A \cup B), x \in \mathcal{R}^3$,
the goal is to find a curve $C$ such that for each point $c \in C$, and its closest projections
\begin{align}
a^c &= \operatornamewithlimits{argmin}\limits_{a \in A} ||a-c|| \text{, and} \\
b^c &= \operatornamewithlimits{argmin}\limits_{b \in B} ||b-c||
\end{align}

@@

\noindent where for each point $r$ on line $L(p,q)$, $||r-p|| = ||r-q||$. Moreover, for
any three consecutive sampled points $c_i$, $c_j$ and $c_k$ on curve $C$, the arc lengths
of the curves between their closest projections are the same:
\begin{equation}
\small
 D^A(a^{c_i},a^{c_j}) + D^B(b^{c_i},b^{c_j}) = D^A(a^{c_j},a^{c_k}) + D^B(b^{c_j},b^{c_k})
\end{equation}
\noindent where $D^Z(x,y)$ is the distance along the curve $Z$ between points $x,y \in Z$.

@@

\section{Input}
The input is the six control points for two curves $A$ and $B$, named $A_0$ to $A_5$ and $B_0$ to $B_5$, such that the first and the last control points are the same.

\section{Overview}
The project is composed of the following three main parts: (1) the representation of the curves,
(2) the computation of the average curve, and (3) the visualization of the different curve properties.
The challenge in the representation is that we want the spline curves to first meet at two end points and then,
to have $C^1$ continuity at the intersection of the splines. For the curve average, we had to ensure
the points satisfied the constraints in Equations 1-3. Lastly, the visualization contained multiple challenges
including parallel transport, circular arc computation and etc.

\section{Curve Representation}
We compose each curve out of piecewise quadratic Hermite and cubic Hermite splines. Each spline is connected at a control point by the same local velocity. This leads to a $\mathcal{C}^1$-smooth curve.
We define the tangent/velocity at control point $A_i$ as $T_i = c*(A_{i+1}-A_{i-1})$. The parameter $c$ defines the curvature or the influence of that velocity. In our experiments, we set $c$ to 0.5.

@@

\subsection{Quadratic Hermite Spline}
Given the points $X_0$ and $X_1$ and the tangent $T_0$, find a quadratic spline $P(t)$ so that the following holds:
\begin{equation}
\small
 P(0)=X_0, P(1)=X_1, P'(0)=T_0
\end{equation}
Solving this system of equation leads to the following formula:
\begin{equation}
\small
 P(t) = (t^2-2t+1)X_0 + (-t^2+2t)X_1 + (t^2-t)T_0
\label{eq:QuadraticHermite}
\end{equation}

\subsection{Cubic Hermite Spline}
Given the points $X_0$ and $X_1$ and the tangents $T_0$ and $T_1$, find a cubic spline $P(t)$ so that the following holds:
\begin{equation}
\small
 P(0)=X_0, P(1)=X_1, P'(0)=T_0, P'(1)=T_0
\end{equation}
This equation is solved similar to the quadratic case, leading to:
\begin{equation}
\small
 P(t) = (2t^3-3t^2+1)X_0 + (t^3-2t^2+t)T_0 + (-2t^3+3t^2)X_1 + (t^3-t^2)T_1
\label{eq:CubicHermite}
\end{equation}

\section{Average Curve Computation}

@@

\section{Visualization and Editing}

@@

As you can see, without geodesic sampling, the distances between the samples of the curve average are almost equispaced, but the closest projections on the control curve, indicated by the circular arcs, vary greatly in the distances. With geodesic sampling turned on, the curve average is sampled in that way that the sum of the distances between two consecutive closest projects on the control curves are constant. However, this leads to large variation of the distances of the curve average samples.

@@

\subsection{Circular arc}
Instead of displaying the closest projections as straight lines, we could also draw them using circular arcs.
\begin{figure}[ht!] 
  \centering
	\begin{tikzpicture}[scale=0.5]
		\coordinate[label=right:P] (P) at (0,0);
		\coordinate[label=left:A] (A) at (-6,-2);
		\coordinate[label=right:B] (B) at (6,-2);
		\coordinate[label=right:C] (C) at (0,-20);
		\draw (P)--(A); %closest projections
		\draw (P)--(B);
		\draw[blue,thick] (6.4,-0.8) coordinate -- (5.6,-3.2) coordinate; %parts of the control curve
		\draw[green,thick] (-6.4,-0.8) coordinate -- (-5.6,-3.2) coordinate;
		\draw (-6.2,-1.4) coordinate -- (-5.6,-1.2) coordinate -- (-5.2,-2.4) coordinate -- (-5.8,-2.6) coordinate; % right angles
		\draw (6.2,-1.4) coordinate -- (5.6,-1.2) coordinate -- (5.2,-2.4) coordinate -- (5.8,-2.6) coordinate;
		\path[blue,thick] (6.4,-0.8) coordinate edge[out=70,in=-120] (6.8,1) coordinate; % control curve
		\path[blue,thick] (5.6,-3.2) coordinate edge[out=250,in=100] (5.8,-5) coordinate;
		\path[green,thick] (-6.4,-0.8) coordinate edge[out=110,in=-120] (-7,1) coordinate;
		\path[green,thick] (-5.6,-3.2) coordinate edge[out=290,in=80] (-5,-5) coordinate;
		\path[black,thick] (0,0) coordinate edge[out=90,in=-70] (-0.2,1) coordinate; % average curve
		\path[black,thick] (0,0) coordinate edge[out=-90,in=110] (0.2,-3) coordinate;
		\draw (A)--(C); %circle sides
		\draw (B)--(C);
		\coordinate[label=left:r] (r) at (-4,-8);
		\draw[red,thick] ([shift=(71.567:19)] 0,-20) arc (71.567:108.435:19); %circular arc
		\tikzset{mypoints/.style={fill=white,draw=black,thick}} %coordinate points
		\fill[mypoints] (P) circle (5.0pt);
		\fill[mypoints] (A) circle (5.0pt);
		\fill[mypoints] (B) circle (5.0pt);
		\fill[mypoints] (C) circle (5.0pt);
	\end{tikzpicture}
	\caption{circular arc}
	\label{fig:CircularArc}
\end{figure}
In Fig.\ref{fig:CircularArc}, you can see parts of the two input curves in green and blue and the curve average in black in the middle. The current trace point is $P$ and the closest projections of $P$ on the control points are $A$ and $B$. \\
We can now construct a unique circle in 3d with center $C$, radius $r$ and normal $N$ that goes through $A$ and $B$ and the tangents at these points are $AP$ and $BP$.

@@

We draw the circular arc by approximating it using piecewise straight tubes. The samples are obtained by rotating the vector $CA$ around $N$ until it becomes $CB$. This is done the same way as in the Swirl-project:
\begin{equation}
\begin{array}{lcl}
 CA^{\circ}(\theta; N) &:=& C + U^{\circ}(\alpha; N) \\
                        & =& C + \cos \alpha U - \sin \alpha (N \times U)\\
 \multicolumn{3}{l}{\text{with } V := CA\angle\underline{N} = (W\bullet\underline{N})\underline{N}} \\
 \multicolumn{3}{l}{\text{and } U = CA-V} \\
 \multicolumn{3}{l}{\theta \text{ ranges from } 0 \text{ to } CA^{\wedge}CB=\cos^{-1}(\underline{CA}\bullet\underline{CB})}
\end{array}
\label{eq:Rot}
\end{equation}

Furthermore, if we sample the circular arcs at some angles and combine the samples with the same angle, we obtain a net, as shown in Fig.\ref{GeodesicSampling}d. If both control curves lie in a plane, this net also lies in the same plane.

>>>>

\section{Objective}
Formally, for two curves $A$ and $B$ defined by the control points $\{A_0 ... A_{n-1}\}$ and $\{B_0 ... B_{n-1}\}$, all points are in $\mathbb{R}^3$, and an interpolation scheme,
the goal is to find a curve $C$ such that for each point $c \in C$, and its closest projections
\begin{align}
a^c &= \operatornamewithlimits{argmin}\limits_{a \in A} ||a-c|| \text{, and} \\
b^c &= \operatornamewithlimits{argmin}\limits_{b \in B} ||b-c||
\end{align}

@@

\noindent where for each point $r$ on line $L(p,q)$, $||r-p|| = ||r-q||$. We call the curves $A$ and $B$ the control curves and $C$ the average curve.
Moreover, for any three consecutive sampled points $c_i$, $c_j$ and $c_k$ on curve $C$, the arc lengths
of the curves between their closest projections are the same:
\begin{equation}
\small
 D^A(a^{c_i},a^{c_j}) + D^B(b^{c_i},b^{c_j}) = D^A(a^{c_j},a^{c_k}) + D^B(b^{c_j},b^{c_k})
\end{equation}
\noindent where $D^Z(x,y)$ is the distance along the curve $Z$ between points $x,y \in Z$.

@@

It can also be seen as a generalization of the medial axis transformation from the 2D case to the 3D case: 
In every projection of the $\mathbb{R}^3$ in the $\mathbb{R}^2$, the average curve $C$ is the medial axis of $A$ and $B$.

\section{Input}
For simplicity, we only focus on a special case that each control curve is defined by six control points $A_0$ to $A_5$ and $B_0$ to $B_5$, such that the first and the last control points are the same.

\section{Overview}
The project is composed of the following three main parts: (1) the representation of the curves (see \ref{chapter:CurveRepr}),
(2) the computation of the average curve (see \ref{chapter:Tracing}), and (3) the visualization of the different curve properties (see \ref{chapter:Visual}).
The challenge in the representation is that we want the spline curves to first meet at two end points and then,
to have $C^1$ continuity at the intersection of the splines. For the curve average, we had to ensure
the points satisfied the constraints in Equations 1-3. Lastly, the visualization contained multiple challenges
including parallel transport, circular arc computation and etc.

\section{Curve Representation}\label{chapter:CurveRepr}
We compose each curve out of piecewise quadratic Hermite and cubic Hermite splines. Each spline is connected at a control point by the same local velocity. This leads to a $\mathcal{C}^1$-smooth curve.
We define the tangent/velocity at control point $A_i$ as $T_i = c(A_{i+1}-A_{i-1})$. The parameter $c$ defines the curvature or the influence of that velocity. In our experiments, we set $c$ to 0.5.

@@

\subsection{Quadratic Hermite Spline}
Given the points $X_0$ and $X_1$ and the tangent $T_0$, find a quadratic spline $P(t)$ so that the following holds:
\begin{equation}
 P(0)=X_0, P(1)=X_1, P'(0)=T_0
\end{equation}
Solving this system of equation leads to the following formula:
\begin{equation}
 P(t) = (t^2-2t+1)X_0 + (-t^2+2t)X_1 + (t^2-t)T_0
\label{eq:QuadraticHermite}
\end{equation}

\subsection{Cubic Hermite Spline}
Given the points $X_0$ and $X_1$ and the tangents $T_0$ and $T_1$, find a cubic spline $P(t)$ so that the following holds:
\begin{equation}
 P(0)=X_0, P(1)=X_1, P'(0)=T_0, P'(1)=T_0
\end{equation}
This equation is solved similar to the quadratic case, leading to:
\begin{equation}
\begin{array}{rcl}
	P(t) &=& (2t^3-3t^2+1)X_0 + (t^3-2t^2+t)T_0 \\
	     &+& (-2t^3+3t^2)X_1 + (t^3-t^2)T_1
\end{array}
\label{eq:CubicHermite}
\end{equation}

\section{Average Curve Computation}\label{chapter:Tracing}

@@

\subsection{Detect non-compatible control curves}
TODO

\section{Visualization and Editing}\label{chapter:Visual}

@@

As you can see, without geodesic sampling, the distances between the samples of the curve average are almost equispaced, but the closest projections on the control curve, indicated by the circular arcs, vary greatly in the distances. With geodesic sampling turned on, the curve average is sampled in that way that the sum of the distances between two consecutive closest projects on the control curves is constant. However, this leads to large variation of the distances of the curve average samples.

@@

\subsection{Circular arc}
Instead of displaying the closest projections as straight lines, we could also draw them using circular arcs.
\begin{figure}[ht!] 
  \centering
	\begin{tikzpicture}[scale=0.5]
		\coordinate[label=right:P] (P) at (0,0);
		\coordinate[label=left:A] (A) at (-6,-2);
		\coordinate[label=right:B] (B) at (6,-2);
		\coordinate[label=right:C] (C) at (0,-20);
		\draw (P)--(A); %closest projections
		\draw (P)--(B);
		\draw[blue,thick] (6.4,-0.8) coordinate -- (5.6,-3.2) coordinate; %parts of the control curve
		\draw[green,thick] (-6.4,-0.8) coordinate -- (-5.6,-3.2) coordinate;
		\draw (-6.2,-1.4) coordinate -- (-5.6,-1.2) coordinate -- (-5.2,-2.4) coordinate -- (-5.8,-2.6) coordinate; % right angles
		\draw (6.2,-1.4) coordinate -- (5.6,-1.2) coordinate -- (5.2,-2.4) coordinate -- (5.8,-2.6) coordinate;
		\path[blue,thick] (6.4,-0.8) coordinate edge[out=70,in=-120] (6.8,1) coordinate; % control curve
		\path[blue,thick] (5.6,-3.2) coordinate edge[out=250,in=100] (5.8,-5) coordinate;
		\path[green,thick] (-6.4,-0.8) coordinate edge[out=110,in=-120] (-7,1) coordinate;
		\path[green,thick] (-5.6,-3.2) coordinate edge[out=290,in=80] (-5,-5) coordinate;
		\path[black,thick] (0,0) coordinate edge[out=90,in=-70] (-0.2,1) coordinate; % average curve
		\path[black,thick] (0,0) coordinate edge[out=-90,in=110] (0.2,-3) coordinate;
		\draw (A)--(C); %circle sides
		\draw (B)--(C);
		\coordinate[label=left:r] (r) at (-4,-8);
		\draw[red,thick] ([shift=(71.567:19)] 0,-20) arc (71.567:108.435:19); %circular arc
		\tikzset{mypoints/.style={fill=white,draw=black,thick}} %coordinate points
		\fill[mypoints] (P) circle (5.0pt);
		\fill[mypoints] (A) circle (5.0pt);
		\fill[mypoints] (B) circle (5.0pt);
		\fill[mypoints] (C) circle (5.0pt);
	\end{tikzpicture}
	\caption{circular arc}
	\label{fig:CircularArc}
\end{figure}
In Fig.\ref{fig:CircularArc}, you can see parts of the two input curves in green and blue and the curve average in black in the middle. The current trace point is $P$ and the closest projections of $P$ on the control points are $A$ and $B$. \\
We can now construct a unique circle in 3d with center $C$ and radius $r$ that goes through $A$ and $B$, the tangents at these points are $AP$ and $BP$ and lies in the convex hull of $A$,$B$ and $P$.

@@

We draw the circular arc by approximating it using piecewise straight tubes. The samples are obtained by rotating the vector $CA$ around $N$ until it becomes $CB$. This is done the same way as in the Swirl-project:
\begin{equation}
\begin{array}{lcl}
 CA^{\circ}(\theta; N) &:=& C + U^{\circ}(\theta; N) \\
                        & =& C + \cos \theta U - \sin \theta (N \times U)\\
 \multicolumn{3}{l}{\text{with } V := CA\angle\underline{N} = (W\bullet\underline{N})\underline{N}} \\
 \multicolumn{3}{l}{\text{and } U = CA-V} \\
 \multicolumn{3}{l}{\theta \text{ ranges from } 0 \text{ to } CA^{\wedge}CB=\cos^{-1}(\underline{CA}\bullet\underline{CB})}
\end{array}
\label{eq:Rot}
\end{equation}

Furthermore, if we sample the circular arcs at some angles and combine the samples with the same angle, we obtain a net, as shown in Fig.\ref{fig:GeodesicSampling}d. If both control curves lie in a plane, this net also lies in the same plane.
